---
title: "Assignment Machine Learning"
author: "Angela Marín"
date: "21 de mayo de 2016"
output: html_document
---

This project predicts the manner six people do barbell lifts. 


##DATA CLEANING

First, the data is uploaded, converting the blank fields into NAs:

```{r}
train=read.csv("pml-training.csv",na.strings=c("NA",""), header=T, sep=",")
test=read.csv("pml-testing.csv",na.strings=c("NA",""), header=T, sep=",")
```

Now, the NAs are analized. The number of NAs in the data are shown below:

```{r}
Empty=sapply(train, function(x){sum(is.na(x))})
table(Empty)
```

As it can be seen, there are 97 columns where all the values are cero. This 97 columns are erased:

```{r}
train1=train[,names(Empty[Empty==0])]
```

The first 7 variables are going to be removed because they are related to the participants and not needed to make the model:

```{r}
train1=train1[,-c(1,2)]
```


##PREDICTION

###Libraries

```{r}
library(caret)
library(rpart)
library(rattle)
```

###Partitioning the data

The idea is to partiotionate the train data to have a new training a nd testing set to be abel to study the accuracy before submitting:

```{r}
inTrain <- createDataPartition(train1$classe, p=0.6, list=FALSE)
mytraining <- train1[inTrain, ]
mytesting <- train1[-inTrain, ]
```

###Decision Trees

First, the train function is used with the method rpart as it is shown below:

```{r}
set.seed(1986)
model_rpart=train(classe ~ ., data = mytraining, method="rpart")
fancyRpartPlot(model_rpart$finalModel)
```

####CrossValidation


```{r}
prediction_rpart <- predict(model_rpart, mytesting)
confusionMatrix(prediction_rpart, mytesting$classe)
```

The accuaracy of this model is 0.49, and the p-value shows a significant model.

The second model, it is calculated with the rpart function, with the class method: 


```{r}
set.seed(1987)
model_class <- rpart(classe ~ ., data=mytraining, method="class")
```

####CrossValidation

```{r}
prediction_class <- predict(model_class, mytesting, type = "class")
confusionMatrix(prediction_class, mytesting$classe)
```

The accuaracy of this model is 0.85, higher than the first model.

###Boosted Regression

The third model is calculated with the gbm method of the train function:

```{r}
#set.seed(1988)
#fitControl <- trainControl(method = "repeatedcv",number = 5, repeats = 1)
#model_gbm=train(classe ~ ., data = mytraining,method="gbm", trControl = fitControl,verbose = FALSE)
```

####CrossValidation

```{r}
#prediction_class <- predict(model_rpart, mytesting, type = "class")
#confusionMatrix(predictionsA1, mytesting$classe)
```


##PREDICTING RESULTS WITH THE TEST DATA

The prediction is going to be made with the second model:

```{r}
prediction <- predict(model_class, test, type = "class")
prediction
```
